### 2021-09-15 analysis/paper/paper.Rmd
update paper, PCA, hclust + dendro

---
title: "SI: Settlements"
author:
  - Petr Pajdla:
      email: petr.pajdla@protonmail.com
      institute: [UofO]
      correspondence: false
  - František Trampota:
      email: trampota@phil.muni.cz
      institute: [UofO]
      correspondence: true
institute:
  - UofO: Department of Archaeology and Museology, Masaryk University, Brno
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      toc: true
      toc_depth: 3
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl"
# abstract: |
keywords: |
  keyword 1; keyword 2; keyword 3
editor_options: 
  markdown: 
    wrap: 80
---

Keywords: `r rmarkdown::metadata$keywords`

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = FALSE,
  comment = "#>",
  # fig.path = "../figures/",
  dpi = 300
)
```

```{r packages, cache=FALSE}
library(sf)
library(tidyverse)
library(spatstat)
library(patchwork)
devtools::load_all(".")
```

# Introduction

The analysis is divided into two scales:

-   Pottery traditions
-   Pottery groups

The study area is broadly defined as *Easter part of Bohemia* (abbreviated *B*)
and *Morava river catchment* (*M*). It spans across the area of the Czech
Republic and bordering regions of Austria and Slovakia.

<!-- ======================================================================= -->

# Background

```{r get-data, cache = FALSE}
derived_data <- "analysis/data/derived_data"
path_maps <- here::here(derived_data, "maps/")
path_data <- here::here(derived_data)
path_temp <- here::here(derived_data, "temp/")

# mask
mask <- st_read(paste0(path_maps, "mask.geojson"), quiet = TRUE)
bbox <- st_read(paste0(path_maps, "bbox.geojson"), quiet = TRUE)

# rivers + admin for plots
admin0 <- st_read(paste0(path_maps, "admin0.geojson"), quiet = TRUE)
admin1 <- st_read(paste0(path_maps, "admin1.geojson"), quiet = TRUE)
rivers0 <- st_read(paste0(path_maps, "rivers0.geojson"), quiet = TRUE)
rivers1 <- st_read(paste0(path_maps, "rivers1.geojson"), quiet = TRUE)

# labels
labs_chrono <- read_rds(here::here(derived_data, "chrono_labels.RDS"))

# database
set_base <- read_rds(here::here(derived_data, "settlements.RDS"))

set_spat <- st_read(here::here(derived_data, "settlements_sf.geojson"), 
                    quiet = TRUE)

settlements1 <- set_spat %>%
  right_join(set_base$period1, by = "id") %>% 
  mutate(label = labs_chrono$chrono1[chrono])
settlements2 <- set_spat %>%
  right_join(set_base$period2, by = "id") %>% 
  mutate(label = labs_chrono$chrono2[chrono])

# colors
mycol <- c("B" = "#7FC97F", "M" = "#BEAED4", "b" = "#7FC97F", "m" = "#BEAED4"
)
```

## Pottery traditions

```{r tab-set1-count}
set_base$period1 %>%
  table_nr_set()
```

```{r plot-set1-count, fig.cap = 'Settlements classified by general pottery traditions across time periods.'}
settlements1 %>%
  plot_nr_set()
```

<!-- ----------------------------------------------------------------------- -->

## Pottery groups

```{r tab-set2-count}
set_base$period2 %>%
  table_nr_set()
```

```{r plot-set2-count, fig.cap = 'Settlements classified by pottery groups across time periods.'}
settlements2 %>%
  plot_nr_set()
```

<!-- ======================================================================= -->

# Settlement density

Settlement density is estimated using KDE. Kernel size of 4 km is used. This
value is completely arbitrary but is selected because settlement clusters are
nicely highlighted at this scale.

## Pottery traditions

```{r set1-kde}
res1_long <- settlements1 %>% 
  est_kde()
```

```{r plot-set1-kde, fig.cap = 'Kernel density estimation for different periods.'}
settlements1 %>% 
  plot_kde()
```

<!-- ----------------------------------------------------------------------- -->

## Pottery groups

```{r set2-kde}
res2_long <- settlements2 %>% 
  est_kde()
```

```{r plot-set2-kde, fig.cap = 'Kernel density estimation for different periods.'}
settlements2 %>% 
  plot_kde()
```

<!-- ======================================================================= -->

# Distance to raw material sources

## Raw materials distribution

Raw material sources are defined as either points or lines. Each raw material
can be identified by single or multiple simple features. If the location of
source used in the Neolithic period is unknown or several sources or even modes
of procurement (e. g. from river bed etc.) are possible, the most probable are
included.

Minimum, mean, median and maximum distances are calculated, median is used 
as a result.

```{r get_data-rm}
rm_pts <- st_read(here::here(derived_data, "rm_points.geojson"), quiet = TRUE) %>% 
  add_rm_type()

rm_lns <- st_read(here::here(derived_data, "rm_lines.geojson"), quiet = TRUE) %>% 
  add_rm_type()
```

```{r plot-rm, fig.cap = "Distribution of raw material sources."}
ggplot() + 
  geom_sf(data = mask, fill = "gray80", color = NA, alpha = 0.4) +
  geom_sf(data = rivers1, color = "gray60", size = 0.4, alpha = 0.4) +
  geom_sf(data = rivers0, color = "gray60", size = 0.6, alpha = 0.4) +
  geom_sf(data = set_spat, alpha = 0.1, color = "gray80") +
  geom_sf(data = rm_pts, aes(color = rm), shape = 4) +
  geom_sf(data = rm_lns, aes(color = rm)) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  facet_wrap(vars(label)) +
  theme_void() +
  theme(panel.border = element_rect(color = "black", fill = NA)) +
  labs(color = "Raw material")
```

## Distances

```{r rm-dist, cache = FALSE}
# calculated separately in a script `rm_distance.R`
# dist_pts <- dist_to_rm(set_spat, rm_pts, "rm")
# dist_lns <- dist_to_rm(set_spat, rm_lns, "rm")

rm_dist <- read_csv(paste0(path_temp, "rm_dist.csv")) %>% 
  mutate(rm = str_to_lower(rm))

# rm_dist_mean <- rm_dist %>%
#   select(-region, -type, -shp, -min, -median, -max) %>%
#   group_by(id, rm) %>%
#   summarise(dist_rm = mean(mean)) %>%
#   pivot_longer(cols = dist_rm, names_to = "variable") %>%
#   mutate(variable = str_c(variable, rm, sep = "_")) %>%
#   select(-rm)
```

<!-- ```{r plot-rm-dist-min, fig.cap="Min. distance to raw material sources."} -->
<!-- rm_dist %>% -->
<!--   ggplot(aes(x = min, y = rm, fill = region)) + -->
<!--   geom_violin() + -->
<!--   scale_fill_brewer(palette = "Accent") + -->
<!--   facet_wrap(~type, scales = "free_y") + -->
<!--   labs(x = "min. dist. (km)", y = "raw material", fill = "Region") + -->
<!--   theme_minimal() + -->
<!--   theme(legend.position = "bottom",  -->
<!--         panel.border = element_rect(color = "black", fill = NA)) -->
<!-- ``` -->

## Settlement density against distance to raw materials

> Hypothesis: In Morava river catchment settlement position is directly associated with 
raw material source.

We compare settlement density (KDE) with distance to chosen raw materials.
There is no strong evidence that the distribution of settlements is associated with distance to 
any of the key raw material sources occurring in the region in any of the studied periods.
There are some mild negative correlations in the period 4.8 -- 4.2 k BCE for polished stone tools materials.
Strongest connection is observed between given pottery groups in period 4.8 -- 4.6 k BCE in case of ad/pmd/zelesice raw material sources.

```{r plot-rm-dist-kde-chrono1, fig.cap="KDE against distance to raw material source, Morava river catchment, pottery traditions."}
res1_long %>% plot_dist_lm()
```

```{r tab-rm-dist-cor-kde1}
res1_long %>% tab_dist_lm()
```

```{r plot-rm-dist-kde-chrono2, fig.cap="KDE against distance to raw material source, Morava river catchment, pottery groups."}
res2_long %>% plot_dist_lm()
```

```{r tab-rm-dist-cor-kde2}
res2_long %>% tab_dist_lm()
```

<!--
## PCA for raw material distances

Principal components analysis is used to reduce dimensionality.
First three PCs are enough to cover more than 95% variability in the data set.

```{r rm-dist-pca}
# scale_col <- function(x){
#   (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
# }
# 
# rm_dist_mean_wide <- rm_dist_mean %>%
#   mutate(variable = str_remove(variable, "dist_rm_")) %>% 
#   pivot_wider(names_from = variable, values_from = value) %>% 
#   ungroup() %>% 
#   mutate(across(where(is.numeric), scale_col))
# 
# # matrix for pca to reduce dimensions
# rm_dist_mx <- rm_dist_mean_wide %>% 
#   select(-id) %>% 
#   as.matrix() 
# row.names(rm_dist_mx) <- rm_dist_mean_wide$id
# 
# rm_pca <- prcomp(rm_dist_mx)
# summary(rm_pca)
```

```{r rm-corrplot, fig.cap="Correlation between principal components and given raw materials."}
# corrplot::corrplot(t(rm_pca$rotation[, 1:3]))
```
```{r rm-biplot, fig.cap="First thee principal components (green - East part of Bohemia, blue - Morava river catchment)."}
# pairs(rm_pca$x[, 1:3], col = unname(mycol[str_extract(rownames(rm_dist_mx), ".")]))
```

```{r update-result-rm}
# rm_dist_pca <- rm_pca$x[, 1:3] %>% 
#   as_tibble(rownames = "id") %>% 
#   rename("rm_pc1" = "PC1",
#          "rm_pc2" = "PC2",
#          "rm_pc3" = "PC3",) %>% 
#   pivot_longer(-id, "variable", "value")
# 
# # update result tibbles
# res1_long <- set_base$period1 %>% 
#   select(id, period, chrono) %>% 
#   inner_join(rm_dist_pca) %>%
#   bind_rows(res1_long)
# 
# res2_long <- set_base$period2 %>% 
#   select(id, period, chrono) %>% 
#   inner_join(rm_dist_pca) %>%
#   bind_rows(res2_long)
# 
# rm(list = ls(pattern = "^rm_dist_"))
```
-->

## RM relevant to given period and/or region

```{r rm-period}
# 0 = not used
# 0.5 = somewhat used
# 1 = used
rm_b <- readxl::read_xlsx(here::here("analysis/data/raw_data",
                                     "sidliste_vs_zdroje.xlsx"), sheet = "B")
rm_m <- readxl::read_xlsx(here::here("analysis/data/raw_data",
                                     "sidliste_vs_zdroje.xlsx"), sheet = "M")

rm_relevance <- bind_rows(b = rm_b, m = rm_m, .id = "reg") %>% 
  mutate(across(starts_with("p"), function(x) x * 0.5)) %>% 
  pivot_longer(starts_with("p"), 
               names_to = "period", values_to = "relevance") %>% 
  rename(rm = surovina)

rm_dist_rel_p1 <- set_base$period2 %>% 
  filter_rel_dist_rm()

rm_dist_rel_p2 <- set_base$period2 %>% 
  filter_rel_dist_rm()
```

### RM relations in *general* pottery traditions

#### Morava river catchment

```{r rm-dist-m-p1, fig.cap="Distance to RM sources, Morava river catchment, pottery traditions. RM sources relevant in given period in black, somewhat relevant RM sources in gray."}
rm_dist_rel_p1 %>% 
  plot_rm_reg_dist("m")
```

#### Eastern part of Bohemia

```{r rm-dist-b-p1, fig.cap="Distance to RM sources, Eastern part of Bohemia, pottery traditions. RM sources relevant in given period in black, somewhat relevant RM sources in gray."}
rm_dist_rel_p1 %>% 
  plot_rm_reg_dist("b")
```

### RM relations in *detailed* pottery groups

#### Morava river catchment

```{r rm-dist-m-p2, fig.cap="Distance to RM sources, Morava river catchment, pottery groups. RM sources relevant in given period in black, somewhat relevant RM sources in gray."}
rm_dist_rel_p2 %>% 
  plot_rm_reg_dist("m")
```

#### Eastern part of Bohemia

```{r rm-dist-b-p2, fig.cap="Distance to RM sources, Eastern part of Bohemia, pottery groups. RM sources relevant in given period in black, somewhat relevant RM sources in gray."}
rm_dist_rel_p2 %>% 
  plot_rm_reg_dist("b")
```

```{r rm-dist-rel}
rm_types <- rm_dist %>% 
  select(rm, type) %>% 
  distinct()

rm_types_tab <- rm_types %>% 
  pull(type)

names(rm_types_tab) <- rm_types %>% 
  pull(rm)

res1_long <- rm_dist_rel_p1 %>% 
  prep_dist_rel_rm() %>% 
  bind_rows(res1_long)

res2_long <- rm_dist_rel_p2 %>% 
  prep_dist_rel_rm() %>% 
  bind_rows(res2_long)
```

```{r rm-dist-rel1, fig.cap="Mean distance to relevant raw materials in pottery traditions."}
rm_dist_rel_p1 %>% 
  count_dist_rel_rm() %>% 
  plot_dist_rel_rm()
```

```{r rm-dist-rel2, fig.cap="Mean distance to relevant raw materials in pottery groups."}
rm_dist_rel_p2 %>% 
  count_dist_rel_rm() %>% 
  plot_dist_rel_rm()
```

<!-- ======================================================================= -->

# Density of watercourses

In settlement structure studies it is often assumed that past human settlements
have some kind of a relation to the network of watercourses. Numerous problems
are faced while assessing the relationship between settlements and close
watercourse. Most importantly the man-made recent changes in the structure and
intensity of watercourses (and landscape in general) are a major factor
impairing such analysis.

Usually, the settlement - watercourse relationship is expressed in some form of
a metric giving shortest distance or so. We choose an approach used to study
road networks in modern urban areas [e. g. @lin2020].

Most of the input data is derived from *DIBAVOD* data set A01 layer
(<https://www.dibavod.cz/>).

Density (kernel smoothed intensity) is estimated using `spatstat::density.psp`
function with sigma 2000 m.

```{r get-raster}
# pkg raster
# hydro_dens <- raster::raster(
#   here::here(derived_data, "hydro", "density_raster.tif"))
# 
# hydro_dens_at_points <- hydro_dens %>% 
#   raster::extract(set_spat) %>% 
#   as_tibble() %>% 
#   rename(hydro_kde = value) %>% 
#   bind_cols(select(st_drop_geometry(set_spat), id))

# pkg stars
hydro_dens <- stars::read_stars(
  here::here(derived_data, "hydro", "density_raster.tif")
) %>% 
  st_set_crs(value = st_crs(set_spat))

hydro_dens_at_points <- hydro_dens %>% 
  stars::st_extract(set_spat) %>% 
  st_drop_geometry() %>% 
  rename(hydro_kde = density_raster.tif) %>% 
  bind_cols(select(st_drop_geometry(set_spat), id))

# bind to results
res1_long <- set_base$period1 %>% 
  prep_hydro_dens_res() %>% 
  bind_rows(res1_long)

res2_long <- set_base$period2 %>% 
  select(-facet) %>% 
  prep_hydro_dens_res() %>% 
  bind_rows(res2_long)
```

```{r plot-hydro-density, fig.cap="Kernel smoothed intensity of watercourses overlaid with sites (raster clipped to research area extent)."}
# hydro_dens %>% 
#   raster::mask(mask) %>% 
#   raster::rasterToPoints() %>%
#   as_tibble() %>%
#   rename(KDE = density_raster) %>% 
#   ggplot() +
#   geom_raster(aes(x, y, fill = KDE)) +
#   geom_sf(data = set_spat, alpha = 0.1, shape = 1, color = "white") +
#   theme_void()

ggplot() +
  stars::geom_stars(data = st_crop(hydro_dens, mask), sf = TRUE) +
  geom_sf(data = set_spat, alpha = 0.1, shape = 1, color = "white") +
  theme_void() +
  labs(fill = "KDE")
```
<!--- ====================================================================== -->

# Terrain surrounding the settlements

We focus on slope around the settlements.

```{r dem-data}
dem_data <- read_csv(here::here(derived_data, "dem/vals_dem.csv"))

# bind to results
res1_long <- set_base$period1 %>% 
  prep_dem_res() %>% 
  bind_rows(res1_long)

res2_long <- set_base$period2 %>% 
  select(-facet) %>% 
  prep_dem_res() %>% 
  bind_rows(res2_long)
```

```{r altitude-to-res}
altitude <- set_spat %>% st_drop_geometry() %>% 
  as_tibble() %>% 
  select(id, altitude)

# bind to results
res1_long <- set_base$period1 %>% 
  prep_alt_res() %>% 
  bind_rows(res1_long)

res2_long <- set_base$period2 %>% 
  select(-facet) %>% 
  prep_alt_res() %>% 
  bind_rows(res2_long)

```

```{r dem-analysis-correlation, fig.cap='Correlation between altitude derived from DEM and altitude reading in the topographic map.'}
alt_plot <- set_spat %>% st_drop_geometry() %>% 
  as_tibble() %>% 
  select(id, altitude) %>% 
  left_join(dem_data)

alt_cor <- alt_plot %>% 
  summarise(cor = cor(altitude, dem)) %>% 
  pull(cor) %>% 
  round(2)

alt_plot %>% 
  ggplot(aes(altitude, dem)) +
  geom_point() +
  theme_minimal() +
  annotate(geom = "text", x = 175, y = 525, 
           label = paste("Cor. ", alt_cor))
```

<!--- ====================================================================== -->

# Settlement 'hierarchy'

Minimal distance to fortified settlements in given regions is explored.

```{r dist-fenced-input}
dist_fenced <- read_rds(here::here(derived_data, "fortif/distance_fortif.RDS"))

# results 1
res1_long <- dist_fenced$traditions %>% 
  prep_dist_fenced() %>% 
  bind_rows(res1_long)

# results 2
dist_fenced_p2 <- dist_fenced$groups %>% 
  prep_dist_fenced() %>% 
  bind_rows(res2_long)
```

<!--- ====================================================================== -->

# Settlement 'continuity'

Is the same location settled in the previous period?

```{r continuity-calc}
res1_long <- set_base$period1 %>%
  settlement_continuity() %>% 
  bind_rows(res1_long)

res2_long <- set_base$period2 %>%
  settlement_continuity() %>% 
  bind_rows(res2_long)
```

```{r continuity-plot1, fig.cap="Settlements continuity, pottery traditions."}
n_set_per1 <- set_base$period1 %>% 
  add_region() %>% 
  group_by(period_label, reg) %>% 
  summarise(sum = n())

res1_long %>% 
  plot_continuity(y = n_set_per1)
```

```{r continuity-plot2, fig.cap="Settlements continuity, pottery groups."}
n_set_per2 <- set_base$period2 %>% 
  add_region() %>% 
  group_by(period_label, reg) %>% 
  summarise(sum = n())

res2_long %>% 
  plot_continuity(y = n_set_per2)
```

<!--- ====================================================================== -->

# Arrangement along natural lines

Arrangement of settlements along natural 'terrain' lines or along waterways is explored.

```{r along_lines}
along_lines <- read_rds(here::here(derived_data, "lines/linear_arrangement.RDS"))

along_lines1 <- along_lines$traditions %>% 
  mutate(chrono = str_remove(chrono, ".$")) %>% 
  group_by(id, chrono, period, variable) %>% 
  summarise(value = max(value)) %>% 
  ungroup(id, chrono, variable) %>% 
  mutate(value = value / n()) %>% 
  select(-chrono) %>% 
  group_by(variable, period, id) %>% 
  summarise(value = mean(value), .groups = "drop")

along_lines2 <- along_lines$groups %>% 
  distinct(id, chrono, period, variable, value) %>% 
  group_by(period) %>% 
  mutate(value = value / n()) %>% 
  select(-chrono) %>% 
  group_by(variable, period, id) %>% 
  summarise(value = mean(value), .groups = "drop")

# results 1
res1_long <- res1_long %>% 
  bind_rows(along_lines1)

# results 2
res2_long <- res2_long %>% 
  bind_rows(along_lines2)
```

```{r clean}
rm(list = ls(pattern = "^rm"))
rm(list = ls(pattern = "^along"))
rm(list = ls(pattern = "line$"))
rm(list = ls(pattern = "violin$"))
rm(list = ls(pattern = "^n_set"))
rm(list = ls(pattern = "^hydro"))
rm(list = ls(pattern = "fenced$"))
rm(list = ls(pattern = "dem_data"))
rm(list = ls(pattern = "^alt"))
```

<!--- ====================================================================== -->

# Synthesis

```{r res-cleaning}
res1_wide <- res1_long %>% prep_wide_res()
# res1_long <- res1_wide %>% 
#   pivot_longer(-c(id, period), names_to = "variable", values_to = "value")

res2_wide <- res2_long %>% prep_wide_res()
# res2_long <- res2_wide %>% 
#   pivot_longer(-c(id, period), names_to = "variable", values_to = "value")

# # scale cols
# res1_wide <- res1_wide %>% 
#   mutate(across(where(is.numeric), scale))
# 
# res2_wide <- res2_wide %>% 
#   mutate(across(where(is.numeric), scale))
```

## Violin plots

```{r variable-labels}
labs_variables <- c(
  "settlements_kde" = "Settlement density",
  "dist_fenced" = "Min. dist. to fenced set. (km)",
  "line_terrain" = "Along terrain line",
  "line_water" = "Along water line",
  "cont" = "Settlement continuity",
  "hydro_kde" = "Watercourse density",
  # "dem" = "Altitude (DEM)", 
  "altitude" = "Altitude (map)",
  "slope" = "Slope (°)",
  "rm_dist_chipped" = "Distance to chipped RM",
  "rm_dist_polished" = "Distance to polished RM"
  # "rm_pc1" = "Distance to RM sources 1",
  # "rm_pc2" = "Distance to RM sources 2",
  # "rm_pc3" = "Distance to RM sources 3"
) %>% 
  map_chr(str_wrap, 12)
```

```{r syn-p1, fig.height=12}
res1_violin <- res1_long %>% plot_violin()
res1_violin
```

```{r syn-p2, fig.height=12}
res2_violin <- res2_long %>% plot_violin()
res2_violin
```

```{r res-violin-out}
ggsave(here::here("analysis/figures", "res1_violin.png"), res1_violin, 
       width = 9, height = 12)
ggsave(here::here("analysis/figures", "res2_violin.png"), res2_violin, 
       width = 9, height = 12)
```

## Line plots

```{r syn-p1-sum, fig.cap="Summaries for different variables across studied periods. Solid line - mean value, wide transparent line - median value.", fig.height=12}
res1_line <- res1_long %>% plot_lines()
res1_line
```

```{r syn-p2-sum, fig.cap="Summaries for different variables across studied periods. Solid line - mean value, wide transparent line - median value.", fig.height=12}
res2_line <- res2_long %>% plot_lines()
res2_line
```

```{r res-line-out}
ggsave(here::here("analysis/figures", "res1_line.png"), res1_line, 
       width = 7, height = 7)
ggsave(here::here("analysis/figures", "res2_line.png"), res2_line, 
       width = 7, height = 7)
```

<!-- ## PCA -->

<!-- ### Pottery traditions -->

<!-- ```{r pca1-all} -->
<!-- # fit PCA -->
<!-- pca1_all_fit <- res1_wide %>%  -->
<!--   # select(-cont) %>%  -->
<!--   select(where(is.numeric)) %>%  -->
<!--   scale(center = TRUE, scale = TRUE) %>%  -->
<!--   prcomp() -->

<!-- # augmented fitted values -->
<!-- pca1_all_aug <- pca1_all_fit %>%  -->
<!--   broom::augment(res1_wide) %>%  -->
<!--   select(-".rownames") %>%  -->
<!--   rename_with(\(x) str_remove(x, ".fitted")) %>%  -->
<!--   settlements::add_region() %>%  -->
<!--   settlements::add_period_label() %>%  -->
<!--   settlements::add_neo() -->

<!-- # rotation matrix -->
<!-- pca1_all_rot <- pca1_all_fit %>% -->
<!--   broom::tidy(matrix = "rotation") %>% -->
<!--   mutate(PC = str_c("PC", PC)) %>% -->
<!--   pivot_wider(names_from = PC) -->

<!-- # var explained -->
<!-- pca1_all_var_exp <- pca1_all_fit %>%  -->
<!--   broom::augment(res1_wide) %>% -->
<!--   summarise(across(contains("PC"), var)) %>%  -->
<!--   pivot_longer(cols = everything(), names_to = "pc", values_to = "variance") %>%  -->
<!--   mutate(var_exp = variance/sum(variance), -->
<!--          cum_var_exp = cumsum(var_exp), -->
<!--          pc = str_replace(pc, ".fitted", "")) %>%  -->
<!--   select(pc, var_exp, cum_var_exp) -->

<!-- # PC cum var explained over 0.9 -->
<!-- pca1_all_var_exp_over <- pca1_all_var_exp %>%  -->
<!--   mutate(n = !(cum_var_exp >= 0.9)) %>%  -->
<!--   pull(n) %>%  -->
<!--   sum() + 1 -->
<!-- ``` -->

<!-- ```{r pca1-all-var-exp-tab} -->
<!-- knitr::kable(pca1_all_var_exp, col.names = c("Comp.", -->
<!--                                              "Prop. of variance",  -->
<!--                                              "Cumulative prop.")) -->
<!-- ``` -->

<!-- ```{r pca1-all-loadings-plot} -->
<!-- # PCA loadings -->
<!-- pca1_all_loadings <- pca1_all_fit %>%  -->
<!--   broom::tidy(matrix = "rotation") %>%  -->
<!--   mutate(component = str_c("PC", PC), -->
<!--          component = fct_reorder(component, PC)) %>%  -->
<!--   filter(PC <= pca1_all_var_exp_over) %>%  -->
<!--   group_by(PC) %>%  -->
<!--   top_n(8, abs(value)) %>%  -->
<!--   ungroup() %>%  -->
<!--   mutate(cor = if_else(value > 0, "positive", "negative"), -->
<!--          column = tidytext::reorder_within(column, abs(value), component)) -->

<!-- pca1_all_loadings %>%  -->
<!--   ggplot(aes(abs(value), column, fill = cor)) + -->
<!--   geom_col(color = "black") + -->
<!--   scale_fill_manual(values = c("white", "gray80")) + -->
<!--   facet_wrap(vars(component), scales = "free_y") + -->
<!--   tidytext::scale_y_reordered() + -->
<!--   theme_bw() + -->
<!--   labs(x = "Abs. contribution value", y = element_blank(), fill = "Value") -->
<!-- ``` -->

<!-- ```{r plot-pca1-results} -->
<!-- plot_pca1_full <- function(x = "PC1", y = "PC2") { -->

<!--   pcs <- c(x, y) -->

<!--   pca1_all_plot <- pca1_all_aug %>%  -->
<!--     ggplot(aes(!!sym(pcs[1]), !!sym(pcs[2]))) + -->
<!--     geom_point(data = select(pca1_all_aug, -period_label),  -->
<!--                alpha = 0.05, size = 0.4) + -->
<!--     geom_point(aes(color = reg),  -->
<!--                alpha = 0.8, size = 0.5, show.legend = FALSE) + -->
<!--     scale_color_manual(values = mycol[c(1, 2)]) + -->
<!--     facet_grid(vars(period_label), vars(reg)) + -->
<!--     theme_minimal() -->

<!--   pca1_all_rot <-  pca1_all_rot %>% -->
<!--     ggplot(aes(!!sym(pcs[1]), !!sym(pcs[2]))) + -->
<!--     geom_segment(xend = 0, yend = 0, alpha = 0.4) + -->
<!--     geom_text( -->
<!--       aes(label = column), -->
<!--       hjust = 1, nudge_x = -0.02, -->
<!--       color = "#904C2F", -->
<!--       size = 2.6 -->
<!--     ) + -->
<!--     theme_minimal() -->

<!--   pca1_all_load <- pca1_all_loadings %>%  -->
<!--     filter(component %in% pcs) %>%  -->
<!--     ggplot(aes(abs(value), column, fill = cor)) + -->
<!--     geom_col(color = "black", show.legend = FALSE) + -->
<!--     scale_fill_manual(values = c("white", "gray80")) + -->
<!--     facet_wrap(vars(component), scales = "free_y") + -->
<!--     tidytext::scale_y_reordered() + -->
<!--     labs(x = "Abs. contribution value", y = element_blank(), fill = "Value") + -->
<!--     theme_minimal() -->

<!--   pca1_all_plot / (pca1_all_rot | pca1_all_load) +  -->
<!--     plot_layout(heights = c(6, 1)) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r pca1-all-plots-pc1-pc2} -->
<!-- plot_pca1_full() -->
<!-- ggsave(here::here("analysis/figures/pca1_full_pc12.png"), height = 14) -->
<!-- ``` -->

<!-- ```{r pca1-all-plots-pc3-pc4} -->
<!-- plot_pca1_full("PC3", "PC4") -->
<!-- ggsave(here::here("analysis/figures/pca1_full_pc34.png"), height = 14) -->
<!-- ``` -->

<!-- ```{r pca1-all-plots-pc5-pc6} -->
<!-- plot_pca1_full("PC5", "PC6") -->
<!-- ggsave(here::here("analysis/figures/pca1_full_pc56.png"), height = 14) -->
<!-- ``` -->

<!-- ```{r pca1-all-plots-pc7-pc8} -->
<!-- plot_pca1_full("PC7", "PC8") -->
<!-- ggsave(here::here("analysis/figures/pca1_full_pc78.png"), height = 14) -->
<!-- ``` -->

<!-- ### Pottery groups -->

<!-- ```{r pca2-all} -->
<!-- # fit PCA -->
<!-- pca2_all_fit <- res2_wide %>%  -->
<!--   # select(-cont) %>%  -->
<!--   select(where(is.numeric)) %>%  -->
<!--   scale(center = TRUE, scale = TRUE) %>%  -->
<!--   prcomp() -->

<!-- # augmented fitted values -->
<!-- pca2_all_aug <- pca2_all_fit %>%  -->
<!--   broom::augment(res2_wide) %>%  -->
<!--   select(-".rownames") %>%  -->
<!--   rename_with(\(x) str_remove(x, ".fitted")) %>%  -->
<!--   settlements::add_region() %>%  -->
<!--   settlements::add_period_label() %>%  -->
<!--   settlements::add_neo() -->

<!-- # rotation matrix -->
<!-- pca2_all_rot <- pca2_all_fit %>% -->
<!--   broom::tidy(matrix = "rotation") %>% -->
<!--   mutate(PC = str_c("PC", PC)) %>% -->
<!--   pivot_wider(names_from = PC) -->

<!-- # var explained -->
<!-- pca2_all_var_exp <- pca2_all_fit %>%  -->
<!--   broom::augment(res2_wide) %>% -->
<!--   summarise(across(contains("PC"), var)) %>%  -->
<!--   pivot_longer(cols = everything(), names_to = "pc", values_to = "variance") %>%  -->
<!--   mutate(var_exp = variance/sum(variance), -->
<!--          cum_var_exp = cumsum(var_exp), -->
<!--          pc = str_replace(pc, ".fitted", "")) %>%  -->
<!--   select(pc, var_exp, cum_var_exp) -->

<!-- # PC cum var explained over 0.9 -->
<!-- pca2_all_var_exp_over <- pca2_all_var_exp %>%  -->
<!--   mutate(n = !(cum_var_exp >= 0.9)) %>%  -->
<!--   pull(n) %>%  -->
<!--   sum() + 1 -->
<!-- ``` -->

<!-- <!-- -->
<!-- ```{r} -->
<!-- fuu <- res2_wide %>% mutate(across(where(is.numeric), \(x) x == 0)) %>%  -->
<!-- select(where(is.logical)) %>%  -->
<!-- rowSums() %>%  -->
<!-- as_tibble() -->

<!-- shit <- res2_wide %>% select(id, period) %>% bind_cols(fuu) %>% arrange(desc(value), id, period) %>%  -->
<!-- filter(value == 8) %>% pull(id) %>% factor() %>% levels() -->

<!-- settlements2 %>%  -->
<!-- filter(id %in% shit) %>%  -->
<!-- sf::st_transform(4326) %>%  -->
<!-- leaflet::leaflet() %>%  -->
<!-- leaflet::addTiles() %>%  -->
<!-- leaflet::addMarkers() -->

<!-- res2_wide %>% filter(id %in% shit) %>% View() -->
<!-- ``` -->
<!-- -->

<!-- ```{r pca2-all-var-exp-tab} -->
<!-- knitr::kable(pca1_all_var_exp, col.names = c("Comp.", -->
<!--                                              "Prop. of variance",  -->
<!--                                              "Cumulative prop.")) -->
<!-- ``` -->

<!-- ```{r pca2-all-loadings-plot} -->
<!-- # PCA loadings -->
<!-- pca2_all_loadings <- pca2_all_fit %>%  -->
<!--   broom::tidy(matrix = "rotation") %>%  -->
<!--   mutate(component = str_c("PC", PC), -->
<!--          component = fct_reorder(component, PC)) %>%  -->
<!--   filter(PC <= pca2_all_var_exp_over) %>%  -->
<!--   group_by(PC) %>%  -->
<!--   top_n(8, abs(value)) %>%  -->
<!--   ungroup() %>%  -->
<!--   mutate(cor = if_else(value > 0, "positive", "negative"), -->
<!--          column = tidytext::reorder_within(column, abs(value), component)) -->

<!-- pca2_all_loadings %>%  -->
<!--   ggplot(aes(abs(value), column, fill = cor)) + -->
<!--   geom_col(color = "black") + -->
<!--   scale_fill_manual(values = c("white", "gray80")) + -->
<!--   facet_wrap(vars(component), scales = "free_y") + -->
<!--   tidytext::scale_y_reordered() + -->
<!--   theme_bw() + -->
<!--   labs(x = "Abs. contribution value", y = element_blank(), fill = "Value") -->
<!-- ``` -->

<!-- ```{r plot-pca2-results} -->
<!-- plot_pca2_full <- function(x = "PC1", y = "PC2") { -->

<!--   pcs <- c(x, y) -->

<!--   pca2_all_plot <- pca2_all_aug %>%  -->
<!--     ggplot(aes(!!sym(pcs[1]), !!sym(pcs[2]))) + -->
<!--     geom_point(data = select(pca2_all_aug, -period_label),  -->
<!--                alpha = 0.05, size = 0.4) + -->
<!--     geom_point(aes(color = reg),  -->
<!--                alpha = 0.8, size = 0.5, show.legend = FALSE) + -->
<!--     scale_color_manual(values = mycol[c(1, 2)]) + -->
<!--     facet_grid(vars(period_label), vars(reg)) + -->
<!--     theme_minimal() -->

<!--   pca2_all_rot <-  pca2_all_rot %>% -->
<!--     ggplot(aes(!!sym(pcs[1]), !!sym(pcs[2]))) + -->
<!--     geom_segment(xend = 0, yend = 0, alpha = 0.4) + -->
<!--     geom_text( -->
<!--       aes(label = column), -->
<!--       hjust = 1, nudge_x = -0.02, -->
<!--       color = "#904C2F", -->
<!--       size = 2.6 -->
<!--     ) + -->
<!--     theme_minimal() -->

<!--   pca2_all_load <- pca2_all_loadings %>%  -->
<!--     filter(component %in% pcs) %>%  -->
<!--     ggplot(aes(abs(value), column, fill = cor)) + -->
<!--     geom_col(color = "black", show.legend = FALSE) + -->
<!--     scale_fill_manual(values = c("white", "gray80")) + -->
<!--     facet_wrap(vars(component), scales = "free_y") + -->
<!--     tidytext::scale_y_reordered() + -->
<!--     labs(x = "Abs. contribution value", y = element_blank(), fill = "Value") + -->
<!--     theme_minimal() -->

<!--   pca2_all_plot / (pca2_all_rot | pca2_all_load) +  -->
<!--     plot_layout(heights = c(6, 1)) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r pca2-all-plots-pc1-pc2} -->
<!-- plot_pca2_full() -->
<!-- ggsave(here::here("analysis/figures/pca2_full_pc12.png"), height = 14) -->
<!-- ``` -->

<!-- ```{r pca2-all-plots-pc3-pc4} -->
<!-- plot_pca2_full("PC3", "PC4") -->
<!-- ggsave(here::here("analysis/figures/pca2_full_pc34.png"), height = 14) -->
<!-- ``` -->

<!-- ```{r pca2-all-plots-pc5-pc6} -->
<!-- plot_pca2_full("PC5", "PC6") -->
<!-- ggsave(here::here("analysis/figures/pca2_full_pc56.png"), height = 14) -->
<!-- ``` -->

<!-- ```{r pca2-all-plots-pc7-pc8} -->
<!-- plot_pca2_full("PC7", "PC8") -->
<!-- ggsave(here::here("analysis/figures/pca2_full_pc78.png"), height = 14) -->
<!-- ``` -->

## Principal components analysis

```{r nested-pca}
# traditions
# traditions, by periods
pca1_per <- res1_wide %>%
  add_region() %>% 
  group_by(reg, period) %>% 
  nest() %>% 
  nested_pca()

# traditions, by neo BC
pca1_neo <- res1_wide %>%
  add_region() %>% 
  add_neo() %>% 
  select(-period) %>% 
  group_by(reg, neo) %>% 
  nest() %>% 
  nested_pca()

# groups
# groups, by periods
pca2_per <- res2_wide %>%
  add_region() %>% 
  group_by(reg, period) %>% 
  nest() %>% 
  nested_pca()

# groups, by neo BC
pca2_neo <- res2_wide %>%
  add_region() %>% 
  add_neo() %>% 
  select(-period) %>% 
  group_by(reg, neo) %>% 
  nest() %>% 
  nested_pca()

# pca1_neo %>%
#   select(reg, neo, sdev) %>%
#   unnest(sdev) %>%
#   ungroup(neo, reg) %>%
#   ggplot(aes(factor(pc), cum_prop)) +
#   geom_point() +
#   geom_path(aes(group = neo)) +
#   geom_hline(yintercept = 0.9) +
#   facet_grid(vars(neo), vars(reg)) +
#   scale_y_continuous(limits = c(0, 1)) +
#   theme_minimal()
```

Variable labels in PCA plots:

altitude - alt.,
cont - cont.,
dist_fenced - d. fenced,
hydro_kde - hydro.,
line_terrain -  t. line,
settlements_kde - dens.,
slope ~ slope,
rm_dist_chipped - d. ch.,
line_water - w. line,
rm_dist_polished - d. pol.

### Pottery traditions in Neolithic B and C

```{r plot-pca1-neo1, fig.cap="PCA for pottery traditions data in Neolithic B and C, PC 1-2 space."}
pca1_neo %>% plot_nested_pca("neo")
```

```{r plot-pca1-neo2, fig.cap="PCA for pottery traditions data in Neolithic B and C, PC 3-4 space."}
pca1_neo %>% plot_nested_pca("neo", c("PC3", "PC4"))
```

### Pottery groups in Neolithic B vs Neolithic C

```{r plot-pca2-neo1, fig.cap="PCA for pottery groups data in Neolithic B and C, PC 1-2 space."}
pca2_neo %>% plot_nested_pca("neo")
```

```{r plot-pca2-neo2, fig.cap="PCA for pottery groups data in Neolithic B and C, PC 3-4 space."}
pca2_neo %>% plot_nested_pca("neo", c("PC3", "PC4"))
```

### Pottery traditions across periods

```{r plot-pca1-period1, fig.cap="PCA for pottery traditions data across periods, PC 1-2 space."}
pca1_per %>% 
  add_period_label() %>% 
  plot_nested_pca("period_label")
```

```{r plot-pca1-period2, fig.cap="PCA for pottery traditions data across periods, PC 3-4 space."}
pca1_per %>% 
  add_period_label() %>% 
  plot_nested_pca("period_label", c("PC3", "PC4"))
```

### Pottery groups across periods

```{r plot-pca2-period1, fig.cap="PCA for pottery groups data across periods, PC 1-2 space."}
pca2_per %>% 
  add_period_label() %>% 
  plot_nested_pca("period_label")
```

```{r plot-pca2-period2, fig.cap="PCA for pottery groups data across periods, PC 3-4 space."}
pca2_per %>% 
  add_period_label() %>% 
  plot_nested_pca("period_label", c("PC3", "PC4"))
```

## Variables clustering

### Pottery traditions

```{r pca1-neo-hclust, fig.cap="Variables clustering dendrogram for pottery traditions."}
hclust1_neo <- pca1_neo %>% 
  hclust_variables()

hclust_plot_dendro(hclust1_neo, c("reg", "neo"))
```

### Pottery groups

```{r pca2-neo-hclust, fig.cap="Variables clustering dendrogram for pottery groups."}
hclust2_neo <- pca2_neo %>% 
  hclust_variables()

hclust_plot_dendro(hclust2_neo, c("reg", "neo"))
```

<!--- ====================================================================== -->

# Results

```{r results-export}
res1_long %>%
  distinct(id, period, variable, .keep_all = TRUE) %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  arrange(id, period) %>% 
  write_csv(here::here(derived_data, "results", "results1.csv"))

res2_long %>% 
  distinct(id, period, variable, .keep_all = TRUE) %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  arrange(id, period) %>% 
  write_csv(here::here(derived_data, "results", "results2.csv"))
```

<!--- ====================================================================== -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational
environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```