### 2022-03-23 analysis/groups/groups.Rmd
add new docs edit orig paper

---
title: "SI: Settlements / Pottery groups"
author: Petr Pajdla
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      toc: true
      toc_depth: 3
bibliography: ../paper/references.bib
csl: "../templates/journal-of-archaeological-science.csl"
# abstract: |
# keywords: |
#   keyword 1; keyword 2; keyword 3
editor_options: 
  markdown: 
    wrap: 80
---

<!-- Keywords: `r rmarkdown::metadata$keywords` -->

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = FALSE,
  comment = "#>",
  # fig.path = "../figures/",
  dpi = 300
)
```

```{r packages, cache=FALSE}
library(sf)
library(tidyverse)
library(spatstat)
library(patchwork)
devtools::load_all(".")
```

# Introduction

The analysis is divided into three scales:

-   Neolithic B/C
-   Pottery traditions
-   Pottery groups

**This document describes results for Pottery groups.**

The study area is broadly defined as *Easter part of Bohemia* (abbreviated *B*)
and *Morava river catchment* (*M*). It spans across the area of the Czech
Republic and bordering regions of Austria and Slovakia.

<!-- ======================================================================= -->

## Data

Data is published in the Journal of Open Archaeology Data:

> Pajdla, P. and Trampota, F., 2021. Neolithic Settlements in Central Europe: Data from the Project ‘Lifestyle as an Unintentional Identity in the Neolithic’. Journal of Open Archaeology Data, 9, p.13. DOI: <http://doi.org/10.5334/joad.88>

```{r load-data, cache = FALSE}
source(here::here("analysis/code/data_load.R"))
```

```{r tab-set-nr}
set_base$period2 %>%
  table_nr_set()
```

```{r map-set-nr, fig.cap = 'Settlements across time periods.'}
settlements2 %>%
  plot_nr_set()
```

<!-- ======================================================================= -->

# Analysis

## Settlement density

Settlement density is estimated using KDE. Kernel size of 4 km is used. This
value is completely arbitrary but is selected because settlement clusters are
nicely highlighted at this scale.

```{r map-kde, fig.cap = 'Kernel density estimation for different periods.'}
settlements2 %>% 
  plot_kde()
```

```{r set-kde}
kde <- settlements2 %>% 
  est_kde() %>% 
  mutate(reg = str_extract(id, "^."))
```

```{r plot-kde}
kde %>% ggplot(aes(period, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  labs(y = "KDE")
```

<!--- ====================================================================== -->

## Terrain surrounding the settlements

### Altitude

```{r alt-data}
alt <- settlements2 %>% 
  select(id, altitude) %>% 
  st_drop_geometry() %>% 
  distinct(id, altitude) %>% 
  transmute(id, value = altitude,
            variable = "altitude")
```

```{r alt}
settlements2 %>% 
  mutate(reg = str_extract(id, "^.")) %>% 
  ggplot(aes(period, altitude)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  labs(y = "Altitude (m)")
```

### Slope

```{r slope}
slope <- read_csv(here::here("analysis/data/derived_data", 
                             "slope_at_points.csv")) %>% 
  filter(id %in% settlements2$id) %>% 
  mutate(variable = "slope")
```

```{r plot-slope}
slope %>% 
  left_join(settlements2, by = "id") %>% 
  mutate(reg = str_extract(id, "^.")) %>% 
  ggplot(aes(period, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  labs(y = "Slope (°)")
```

### TPI

> Terrain / Topographic position index

```{r tpi}
tpi <- read_csv(here::here("analysis/data/derived_data", 
                           "tpi_at_points.csv")) %>% 
  filter(id %in% settlements2$id) %>% 
  mutate(variable = "tpi")
```

```{r plot-tpi}
tpi %>% 
  left_join(settlements2, by = "id") %>% 
  mutate(reg = str_extract(id, "^.")) %>% 
  ggplot(aes(period, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  labs(y = "TPI")
```

### Density of water courses

```{r hydro}
hydro <- read_csv(here::here("analysis/data/derived_data", 
                             "hydro_at_points.csv")) %>% 
  filter(id %in% settlements2$id) %>% 
  mutate(variable = "hydro") %>% 
  rename(value = hydro_kde)
```

```{r plot-hydro}
hydro %>% 
  left_join(settlements2, by = "id") %>% 
  mutate(reg = str_extract(id, "^.")) %>% 
  ggplot(aes(period, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  labs(y = "Hydrology KDE")
```

<!--- ====================================================================== -->

<!-- # Settlement 'hierarchy' -->

<!-- Minimal distance to fortified settlements in given regions is explored. -->

<!--- ====================================================================== -->

# Arrangement along natural lines

```{r along-lines}
along_lines <- read_rds(here::here(derived_data, "lines/linear_arrangement.RDS"))

along_lines <- along_lines$groups %>% 
  mutate(chrono = str_remove(chrono, ".$")) %>% 
  group_by(id, chrono, period, variable) %>% 
  summarise(value = max(value)) %>% 
  ungroup(id, chrono, variable) %>% 
  mutate(value = value / n()) %>% 
  select(-chrono) %>% 
  group_by(variable, period, id) %>% 
  summarise(value = mean(value), .groups = "drop")
```

<!--- ====================================================================== -->

# Distance to raw material sources

```{r rm}
# calculated separately in a script `rm_distance.R`
# dist_pts <- dist_to_rm(set_spat, rm_pts, "rm")
# dist_lns <- dist_to_rm(set_spat, rm_lns, "rm")
rm_dist <- read_csv(here::here("analysis/data/input_data", "rm_dist.csv")) %>% 
  mutate(rm = str_to_lower(rm)) %>% 
  filter(id %in% settlements2$id)

# 0 = not used
# 0.5 = somewhat used
# 1 = used
rm_b <- readxl::read_xlsx(here::here("analysis/data/raw_data",
                                     "sidliste_vs_zdroje.xlsx"), sheet = "B")
rm_m <- readxl::read_xlsx(here::here("analysis/data/raw_data",
                                     "sidliste_vs_zdroje.xlsx"), sheet = "M")

rm_relevance <- bind_rows(b = rm_b, m = rm_m, .id = "reg") %>% 
  mutate(across(starts_with("p"), function(x) x * 0.5)) %>% 
  pivot_longer(starts_with("p"), 
               names_to = "period", values_to = "relevance") %>% 
  rename(rm = surovina)

rm_dist_rel <- set_base$period2 %>% 
  filter_rel_dist_rm()

# ---
rm_types <- rm_dist %>% 
  select(rm, type) %>% 
  distinct()
rm_types_tab <- rm_types %>% 
  pull(type)
names(rm_types_tab) <- rm_types %>% 
  pull(rm)
# ---

rm_dist_fin <- rm_dist_rel %>% 
  prep_dist_rel_rm()
```

<!--- ====================================================================== -->

# Synthesis

```{r variable-labels}
labs_variables <- c(
  "settlements_kde" = "Settlement density",
  # "dist_fenced" = "Min. dist. to fenced set. (km)",
  "line_terrain" = "Along terrain line",
  "line_water" = "Along water line",
  # "cont" = "Settlement continuity",
  "hydro" = "Watercourse density",
  # "dem" = "Altitude (DEM)", 
  "altitude" = "Altitude (map)",
  "slope" = "Slope (°)",
  "tpi" = "TPI",
  "rm_dist_chipped" = "Distance to chipped RM",
  "rm_dist_polished" = "Distance to polished RM"
  # "rm_pc1" = "Distance to RM sources 1",
  # "rm_pc2" = "Distance to RM sources 2",
  # "rm_pc3" = "Distance to RM sources 3"
) %>% 
  map_chr(str_wrap, 12)
```

```{r variables}
res_long <- alt %>% 
  bind_rows(slope) %>% 
  bind_rows(tpi) %>% 
  bind_rows(hydro) %>% 
  left_join(set_base$period2) %>% 
  select(id, variable, value, period) %>% 
  bind_rows(along_lines) %>% 
  bind_rows(rm_dist_fin) %>% 
  bind_rows(select(kde, -reg)) %>% 
  filter(id %in% set_base$period2$id) %>% 
  group_by(variable, id, period) %>% 
  summarise(value = mean(value)) %>% 
  ungroup()

res_wide <- res_long %>% 
  pivot_wider(id_cols = c(id, period), 
              names_from = variable, 
              values_from = value, values_fill = 0)

```

## Principal components analysis

```{r nested-pca}
# traditions
pca_per <- res_wide %>%
  add_region() %>% 
  group_by(reg, period) %>% 
  nest() %>% 
  nested_pca()
```

```{r pca-plot}
pca_per %>% plot_nested_pca(var = "period")
```

## Groups in the data

```{r package-mclust}
library(mclust)
```

```{r mclust2-neo}
mb_clust <- pca_per %>% mbc_cluster()

classif <- mb_clust %>%
  dplyr::select(reg, period, classif) %>% 
  tidyr::unnest(c(classif)) %>% 
  dplyr::ungroup(c(reg, period)) %>% 
  dplyr::mutate(g = str_c(reg, g))

set_spat %>% 
  select(id) %>% 
  inner_join(classif, by = "id") %>% 
  ggplot() +
  geom_sf(aes(color = g)) +
  facet_grid(vars(period), vars(g)) +
  theme_void()
```

```{r spat-out}
# set_spat %>% 
#   select(id) %>% 
#   inner_join(classif, by = "id") %>% 
#   write_sf(here::here("analysis/groups/groups.geojson"), delete_layer = TRUE)
```

```{r plot-vars-g}
plot_g_distribution <- function(var) {
  res_long %>% 
    filter(variable == var) %>% 
    inner_join(classif, by = c("id", "period")) %>% 
    add_period_label() %>% 
    ggplot(aes(g, value)) +
    geom_boxplot() +
    facet_wrap(vars(period_label), scales = "free_x") +
    theme_bw() +
    labs(title = var)
}

plot_g_distribution("altitude")
plot_g_distribution("slope")
plot_g_distribution("tpi")

plot_g_distribution("hydro")

plot_g_distribution("line_terrain")
plot_g_distribution("line_water")

plot_g_distribution("rm_dist_chipped")
plot_g_distribution("rm_dist_polished")

plot_g_distribution("settlements_kde")
```

<!--- ====================================================================== -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational
environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```