### 2022-06-23 README.Rmd
update readme, namespace/desc after build

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Spatial analysis of Neolithic settlement patterns in central Europe"
```

# settlements

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/petrpajdla/settlements/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Trampota F., Pajdla P., (2022). _`r Title`_. Journal? <https://doi.org/xxx/xxx>

The data paper accompanying this article:

> Pajdla P.. Trampota F. (2021). _Neolithic Settlements in Central Europe: Data from the Project 'Lifestyle as an Unintentional Identity in the Neolithic'_. Journal of Open Archaeology Data 9, p.13. <http://doi.org/10.5334/joad.88>.

Data is deposited in a Zenodo repository: <https://doi.org/10.5281/zenodo.5653180>.

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): R Markdown source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. Further, the analysis is divided into three 
    parts with different granularity/detail in time.
  - [:file\_folder: neo](/analysis/neo): Analysis for Neolithic B and C periods.
  - [:file\_folder: neo](/analysis/traditions): Analysis for pottery traditions.
  - [:file\_folder: neo](/analysis/groups): Analysis for pottery groups.
  - [:file\_folder: data](/analysis/data): Data used in the analysis. 
    We were not able to publish all our source data due to copyrights etc.
    There is also an individual data paper: <http://doi.org/10.5334/joad.88>.

## How to run in your broswer or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

The simplest way to explore the text, code and data is to click on
[binder](https://mybinder.org/v2/gh/petrpajdla/settlements/master?urlpath=rstudio)
to open an instance of RStudio in your browser, which will have the
compendium files ready to work with. Binder uses rocker-project.org
Docker images to ensure a consistent and reproducible computational
environment. These Docker images can also be used locally.

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.Rmd")` in the R console

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse, please refer to the individidual data publication at <http://doi.org/10.5334/joad.88>.

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.